Towards Verifying the Bitcoin-S Library

### 2019-06-29 paper/paper.tex
more work

In Bitcoin-S there is a class \texttt{Satoshis} representing an amount of bitcoins.

The additions' signature looks like this:
\begin{lstlisting}[style=scala]
  +(c: CurrencyUnit): CurrencyUnit
\end{lstlisting}

\paragraph{Extracting the relevant Code} The relevant code is in two
files: \texttt{CurrencyUnits.scala} and \texttt{NumberType.scala}.

From those files we removed all code that is not needed for our
verification. For example, we removed all number types except for
Int64 (so Int32, UInt64, etc.) because we do not use them. We also
removed the superclasses Factory and NetworkElement of CurrencyUnit
and Number, respectively, because the inherited members are not used by
the relevant code. Also, we removed all binary operations on Number
that are not used, like subtraction and multiplication.

@@

When we run Stainless on this code (without any properties to prove),
it throws the following errors: describe errors: no support for
  abstract types, unsupported arguments for the BigInt constructor,
  unsupported inheritance for objects.}

@@

Before:
\begin{lstlisting}[style=scala]
sealed abstract class CurrencyUnit {
  def satoshis: Satoshis

  def !=(c: CurrencyUnit): Boolean = !(this == c)

  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis

  def +(c: CurrencyUnit): CurrencyUnit = {
    Satoshis(satoshis.underlying + c.satoshis.underlying)
  }

  protected def underlying: Int64
}
\end{lstlisting}

The addition will now look like this:
\begin{lstlisting}[style=scala]
sealed abstract class CurrencyUnit {
  def satoshis: Satoshis

  def !=(c: CurrencyUnit): Boolean = !(this == c)

  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis

  override def +(c: CurrencyUnit): CurrencyUnit = {
    require(c.satoshis == Satoshis.zero)
    Satoshis(satoshis.underlying + c.satoshis.underlying)
  } ensuring(res => res.satoshis == this.satoshis)

  protected def underlying: Int64
}
\end{lstlisting}

@@

However, currently there are some drawbacks.  For example the sbt
plugin does not always report errors.

@@

Why JAR:
sbt-plugin problems (v 0.1 uses older versions (v0.2 not), no output to console see github \#484)
otherwise use sbt-plugin because of library management and replicability (everyone uses the same Stainless version configured in sbt)
no sbt because

-- version conflicts
-- no output maybe \href{https://github.com/epfl-lara/stainless/issues/484}{issue 484 on
  GitHub}

>>>>

It is of course a crucial property we are verifying here: if zero
satoshis were credited to your account, you would not want your
balance to change! It is also the simplest meaningful property to
verify that we can think of. However, the code involved in performing
an addition of two satoshi amounts in bitcoin-s is non-trivial. The
reason for that is a pecularity of consensus code: agreement with the
majority is more important than correctness, whatever correctness
might mean. The most widely used bitcoin implementation by far is the
reference implementation Bitcoin Core, written in C++. For consensus
code, bitcoin-s has little choice but to be in strict agreement with
the reference implementation. To achieve that, it implements C-like
data types in Scala and then implements functionality using those
C-like data types. For example, the Satoshis class, which is used to
represent an amount of satoshis, is implemented using the class Int64
which represents the C-type \texttt{int64\_t}.

\paragraph{Extracting the relevant Code} The relevant code for the
addition of satoshis is in two files: \texttt{CurrencyUnits.scala} and
\texttt{NumberType.scala}.

\begin{figure}
\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced-nocomments/number/NumberType.scala}
  \caption{Extracted Code from NumberType.scala}
  \label{fig:numbertype}
\end{figure}

\begin{figure}
\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced-nocomments/currency/CurrencyUnits.scala}
  \caption{Extracted Code from CurrencyUnits.scala}
  \label{fig:currencyunits}
\end{figure}

From those files we removed all code that is not needed for the
verification of our property.  For example, we removed all number
types except for Int64 (so Int32, UInt64, etc.) because they are not
used. We also removed the superclasses Factory and NetworkElement of
CurrencyUnit and Number, respectively, because the inherited members
are not used. Also, we removed all binary operations on Number that
are not used, like subtraction and multiplication. The extracted code
is shown in Figure~\ref{fig:numbertype} and
Figure~\ref{fig:currencyunits}.

@@

When we run Stainless on this code (without any properties to prove),
it throws the following errors: describe errors: no support for
  abstract types, unsupported arguments for the BigInt constructor,
  unsupported inheritance for objects.

@@

The addition will now look like this:
\begin{lstlisting}[style=scala]
  def +(c: CurrencyUnit): CurrencyUnit = {
    Satoshis(satoshis.underlying + c.satoshis.underlying)
  } ensuring(res =>
    (c.satoshis == Satoshis.zero) ==>
      (res.satoshis == this.satoshis))
\end{lstlisting}

@@

However, currently there are some drawbacks.  For example the sbt
plugin does not always report errors.
use sbt-plugin because of library management and replicability (everyone uses the same Stainless version configured in sbt)

@@

Why JAR:

no sbt because

-- version conflicts, stainless 0.1 uses older sbt version than bitcoin-s (v0.2 not)
-- no output maybe \href{https://github.com/epfl-lara/stainless/issues/484}{issue 484 on
  GitHub}
